---
title: "CodeBook.rmd"
author: "aleksada"
date: "April 13th, 2015"
output:
  html_document:
    keep_md: yes
---

## Project Description
The project is to prepare tidy data from the accelerometers from the Samsung Galaxy S smartphone that can be used for later analysis

##Study design and data processing
1.Merges the training and the test sets to create one data set.
2.Extracts only the measurements on the mean and standard deviation for each measurement. 
3.Uses descriptive activity names to name the activities in the data set
4.Appropriately labels the data set with descriptive variable names. 
5.From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
###Collection of the raw data
I download the data from 
https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip
###Notes on the original (raw) data 
Additional information can be found at 
http://www.insideactivitytracking.com/data-science-activity-tracking-and-the-battle-for-the-worlds-top-sports-brand/

##Creating the tidy datafile

###Guide to create the tidy data file
1. Download the data folder, and set the downloaded location as my R default directory.
2.Use "list.files" function to store the file path of all data, train data, and test data.
##code
```{r}
file<-list.files("UCI HAR Dataset-2",full.names=TRUE)
train<-list.files("UCI HAR Dataset-2/train",full.names=TRUE)
test<-list.files("UCI HAR Dataset-2/test",full.names=TRUE)
```
3.Merge the main data X.train and X.test, named it as "dataset" 
##code
```{r}
X.train=read.table(train[3])
X.test=read.table(test[3])
dataset=rbind(X.train,X.test)
```
4.Match the measurements' names with their indexes, and then extracts only the measurements on the mean and standard deviation. 
##code
```{r}
feature<-read.table(file[3])
index.mean<-subset(feature, grepl("mean()", V2))
index.std<-subset(feature, grepl("std()", V2))
index<-rbind(index.mean,index.std)
newdata<-dataset[,index$V1]
```
5. Match the descriptive activity names with their indexes and name the activities in the data set, I added it as a new column called activities to newdata, and called the merged dataset as newdata1
##code
```{r}
y.train=read.table(train[4])
y.test=read.table(test[4])
y=rbind(y.train,y.test)
y<-as.factor(y$V1)
label=read.table(file[1])
levels(y)<-label$V2
newdata1<-cbind(newdata,y)
names(newdata1)<-c(as.character(index$V2), 'activities') 
```
6.Labels the data set with descriptive variable names. 
##code
```{r}
names(newdata1)<-c(as.character(index$V2), 'activities') 
```
7.Read the subject file from subject.test and subject.train, then merge it.
After that, add another new column called "subject" to newdata1.
At last, use aggregate function to creates a second, independent tidy data set with the average of each variable for each activity and each subject.
##code
```{r}
sub.train=read.table(train[2])
sub.test=read.table(test[2])
sub=rbind(sub.train,sub.test)
names(newdata2)<-c(names(newdata1),"subject")
tiny_data<-aggregate(.~activities+subject,data=newdata2,mean)
```

##Description of the variables in the tiny_data.txt file
The final tiny data has 81 varibles and 180 observations.
First Two columns are 6 activities and 30 subjects, and the other columns are the measurements on the mean and standard deviation for each measurement
## You can view the final data using following Rcode
```{r}
View(tiny_data)
```
###Measurement 1 (repeat this section for all variables in the dataset)
```{r}
head(newtable[,1:3])
          activities subject tBodyAcc-mean()-X
1            WALKING       1         0.2773308
2   WALKING_UPSTAIRS       1         0.2554617
3 WALKING_DOWNSTAIRS       1         0.2891883
4            SITTING       1         0.2612376
5           STANDING       1         0.2789176
6             LAYING       1         0.2215982
```
We can see that, in the tiny data, the name for the first measurement tBodyAcc-mean()-X, which is the same as the name in features.txt, and for each activities for subject 1, we find the average of the tBodyAcc-mean()-X
